\vskip-0.6in Endogenous Politics and the Design of Trade Institutions

### 2016-03-30 to-do.tex
organize comments from feedback.tex into to-do.tex

March 30, 2016
\begin{enumerate}
	\item Motivation
		\begin{itemize}
			\item From Chad: it's missing intersection with reality
				\begin{itemize}
					\item Well articulated question / paradox
				\end{itemize}
		\end{itemize}
	\item More on escape clause / WEC
		\begin{itemize}
			\item It's particularly interesting that in the presence of a shock the governments will use the escape clause as intended but without a shock a lobby group would always `step into the breach' and cause it to be abused.  I would like to see the paper devote more attention to the conditions under which this problem arises and when it does not.
			\item `Withdrawal of equivalent concessions' (WEC).
				\begin{itemize}
					\item Zissimos (2007): a government `chooses the severity of its own punishment' by the extent of its initial deviation.  In that paper, I took WEC for granted and then played out trade liberalization according to its rules.  But this left open the question of why WEC made sense as an approach to punishment.  Now I'm wondering whether your framework offers an answer.
					\item Under WEC, it is worth deviating from the agreement in proportion to $s$ because your partner deviates to the same extent and that keeps the agreement on track; the point first made by Bagwell and Staiger (1990).
					\item But WEC might eliminate the incentive to respond to lobbying pressure $e$ because when your trade partner deviates by WEC this takes away from you exactly (in a symmetrical framework) what you gained from the lobby for implementing a deviation of that size.
					\item I would find an examination of WEC much more compelling that the approach to punishment that you currently discuss on page 20, whereby two bindings are negotiated: one that applies if the trade partner keeps to the agreement and one that applies if they deviate.  As far as I know this latter approach does not mirror at all the institutional features actually set out by the GATT/WTO.
				\end{itemize}
			\item Possible puzzle to explain: WEC has an effect here, and that effect is weakened by the change in EC rules during the Uruguay Round
			\item Read Chad's ``Why Are Safeguards under the WTO So Unpopular?'' (World Trade Review 2002), at \url{C:\Users\Kristy\Dropbox\Research\PE_shocks\Literature\2016}
				\begin{itemize}
					\item Need to understand better what changed, both with safeguard and with other policy instruments
				\end{itemize}
			\item There is this drop off in usage of escape clause (EC) after Uruguay Round (UR) reforms, even though retaliation prohibited for first three years. Why, when retaliation goes \textit{away}?
			\item I'm not sure there's a way to make sense of Ben's point: in my set-up, government doesn't ``get'' something different from lobby than it does directly from the shock. Maybe it should.
				\begin{itemize}
					\item Government feels $\ga$ the same whether it's elevated because of $s$ or because of $e$
					\item What is the neutralizing that happens? Why would the government invoke EC when it knows that WEC is coming anyway? Because it's in another sector where it puts less weight right now?
				\end{itemize}
			\item Need to go to continuous value of EC tariff
				\begin{itemize}
					\item There will be optimal tariff for whatever value of $s$ is realized
						\begin{itemize}
							\item Not sure how fine I want to go with $s$
						\end{itemize}
					\item Lobby will choose its optimal level of $e$. Will this be constrained in any way? Probably not, if there isn't a fixed EC-binding.
				\end{itemize}
			\end{itemize}
	\item One good idea / do too much
	\item BS2005 vs. MRC2007
	\item Structured differently from most papers
	\item Costly state verification
	\item Define $\ga$ before using it
	\item Take out repeats, especially of references to BS2005, rely to heavily on BS2005
	\item Scrub unnecessary references to weak bindings (since strong are now mostly gone)
\end{enumerate}

>>>>

March 30, 2016
\begin{enumerate}
	\item Motivation
		\begin{itemize}
			\item What it's missing: intersection with reality
				\begin{itemize}
					\item Anecdotes
					\item How policy is used
						\begin{itemize}
							\item I thought that escape clause implications \textit{were} this
						\end{itemize}
					\item Something like how GH predicts export taxes but it doesn't work that way in the real world
					\item Well articulated question / paradox
				\end{itemize}
			\item Look at ``Theory of Managed Trade'' model motivation (BS)
			\item Intro doesn't reflect body of paper
			\item What kinds of empirical questions might we ask with this model?
			\item What would be different because we use this model?
			\item Starting from the introduction, I think you make a convincing case that it is important to endogenize the extent of political pressure and I believe this would be THE selling point of the paper. However, you have not completely sold your contribution and its novelty that you bring in the issue of the objective function of the government. Is this aspect crucial for the point you want to make? If yes, it should be integrate with the previous object from the beginning. If not, it should be postponed a bit.
			\item I also find the introduction too long. You could/should split the literature review that begins on page 4 to a separate section. And in any case, I found that you don't engage with the literature in terms of how you contribute/differentiate from it. Again, your main point seems to get lost into the discussion instead of standing out.
			\item In a sort of conclusion, I think that you need to think a bit more about the objective(s) of the paper and how best to achieve them. And I hope that some of the above is useful!
		\end{itemize}
	\item More on escape clause / WEC
		\begin{itemize}
			\item I started to get really excited about the paper around page 19, especially with Result 3.  The part I really liked was when you analysed the exogenous and endogenous effects on gamma simultaneously around page 28.  In fact I was itching to get to that part right from when you discussed it in the introduction.  I also found the results in the static framework reasonably predictable.  So I'm wondering whether you could dispense with most of the static analysis and get into the repeated game more quickly.
				\begin{itemize}
					\item First, the paper makes the extremely important point above almost in passing.  I would be interested to see a more detailed examination of how this works.  It's particularly interesting that in the presence of a shock the governments will use the escape clause as intended but without a shock a lobby group would always `step into the breach' and cause it to be abused.  I would like to see the paper devote more attention to the conditions under which this problem arises and when it does not.
					\item Then the paper begins to address the equally important question of what can be done about this problem, but again stops short of providing a really compelling answer.  I think it would be worth exploring this in detail.  The paper does get to the heart of the issue, which is that `s' can be observed by the DSB but e cannot.  But the paper then moves on to ex-ante lobbying before getting to the bottom of how to design a dispute settlement system that resolves the conflict created by `e vs s'.
				\end{itemize}
			\item `Withdrawal of equivalent concessions' (WEC).
				\begin{itemize}
					\item Zissimos (2007): a government `chooses the severity of its own punishment' by the extent of its initial deviation.  In that paper, I took WEC for granted and then played out trade liberalization according to its rules.  But this left open the question of why WEC made sense as an approach to punishment.  Now I'm wondering whether your framework offers an answer.
					\item Under WEC, it is worth deviating from the agreement in proportion to $s$ because your partner deviates to the same extent and that keeps the agreement on track; the point first made by Bagwell and Staiger (1990).
					\item But WEC might eliminate the incentive to respond to lobbying pressure $e$ because when your trade partner deviates by WEC this takes away from you exactly (in a symmetrical framework) what you gained from the lobby for implementing a deviation of that size.
						\begin{itemize}
							\item \textbf{Think about how WEC may mitigate LOBBY'S incentives through reaction of gov't}
							\item Ben doesn't have any need for escape, so have to shoehorn WEC into this framework
						\end{itemize}
					\item I would find an examination of WEC much more compelling that the approach to punishment that you currently discuss on page 20, whereby two bindings are negotiated: one that applies if the trade partner keeps to the agreement and one that applies if they deviate.  As far as I know this latter approach does not mirror at all the institutional features actually set out by the GATT/WTO.
				\end{itemize}
			\item Possible puzzle to explain: WEC has an effect here, and that effect is weakened by the change in EC rules during the Uruguay Round
			\item Read Chad's ``Why Are Safeguards under the WTO So Unpopular?'' (World Trade Review 2002), at \url{C:\Users\Kristy\Dropbox\Research\PE_shocks\Literature\2016}
				\begin{itemize}
					\item Need to understand better what changed, both with safeguard and with other policy instruments
				\end{itemize}
			\item There is this drop off in usage of escape clause (EC) after Uruguay Round (UR) reforms, even though retaliation prohibited for first three years. Why, when retaliation goes \textit{away}?
				\begin{itemize}
					\item \textbf{perhaps better abel to respond to $\ga(e)$ before dispute settlement?}
				\end{itemize}
			\item I'm not sure there's a way to make sense of Ben's point: in my set-up, government doesn't ``get'' something different from lobby than it does directly from the shock. Maybe it should.
				\begin{itemize}
					\item Government feels $\ga$ the same whether it's elevated because of $s$ or because of $e$
					\item What is the neutralizing that happens? Why would the government invoke EC when it knows that WEC is coming anyway? Because it's in another sector where it puts less weight right now?
				\end{itemize}
			\item Need to go to continuous value of EC tariff
				\begin{itemize}
					\item There will be optimal tariff for whatever value of $s$ is realized
						\begin{itemize}
							\item Not sure how fine I want to go with $s$
						\end{itemize}
					\item Lobby will choose its optimal level of $e$. Will this be constrained in any way? Probably not, if there isn't a fixed EC-binding.
				\end{itemize}
			\end{itemize}
	\item One good idea / do too much
		\begin{itemize}
			\item You can have one good idea in a paper, the rest get lost
			\item At the moment, I find the paper a bit unstructured in that it does a lot but it doesn't sell in a neat and clear way what it is doing.
			\item If above I suggest to add a section, I think that the paper is already too long and divided into too many sections but that goes back to my first comment about identifying the key issue/contribution: endogeneity of political pressure and/or objective function. The analysis is also split between strict and weak commitments. Which scenario clearly develops the point/contribution you want to make? Shouldn't the other one occupy a smaller place (i.e. space) of the paper?)
		\end{itemize}
	\item BS2005 vs. MRC2007
		\begin{itemize}
			\item Dangerous to build on JLS paper; better to connect to bigger cites or will limit impact, where it can be published
				\begin{itemize}
					\item Perhaps higher level model like BS GE model or some alternative approach that is not so rinky dink (my words)
				\end{itemize}
			\item I like the way you use Bagwell and Staiger (2005) as your benchmark.  But I also wondered whether Maggi and Rodriguez-Clare (2007, henceforth MRC) also offered a useful benchmark.  I know you did have comparisons to their paper in many places but it would be useful if this could be more systematic, and if your paper could draw careful parallels to your assumptions and theirs.  To keep things manageable, I wonder whether you could restrict attention to the case of ex post lobbying.  This is where all the action is, right?  In both papers, the government wants to set a tariff cap in order to be able to extract rents from lobbies afterwards.  But your mechanisms seem quite different.  They have capital mobility whereas your factors are specific to each sector.  If capital is highly immobile they have a hold-up problem, which creates an incentive for lobbies to try to prevent too much liberalization under the agreement.  What drives this feature in your model?  Here I think it would be useful to explain carefully the sequence of events in your set-up more carefully.  What is the relationship between lobby effort level and tariff policy?  Is effort fixed first, after which the government sets its tariff?  If so it seems to me that would drive a kind of hold-up because lobbying effort is sunk.  It would be useful to be clearer about this so we can really understand where the results are coming from and how they compare to MRC.  Having done that, you could bring in ex-ante lobbying and treat it the same way. 
			\item In some parts the paper also seem repeating itself: you review Bagwell and Staiger (2005) in Section 2.2 but you provide another summary of the same paper in Section 3. Again, a revised structure may reduce the overlap, make the paper shorter, and be more to the point (and in general it seems to me that you seem to "dependent" on Bagwell and Staiger... (e.g. beginning of Sections 4 and 6).
		\end{itemize}
	\item Structured differently from most papers
		\begin{itemize}
			\item How did Devashish structure his AER paper: the argument, the model?
				\begin{itemize}
					\item This paper is structured differently than the norm---is there a good reason to do it?
					\item Don't want to challenge what people are used to unless I really need to
					\item In general, how do AER papers \textit{look}?
				\end{itemize}
			\item If above I suggest to add a section, I think that the paper is already too long and divided into too many sections but that goes back to my first comment about identifying the key issue/contribution: endogeneity of political pressure and/or objective function. The analysis is also split between strict and weak commitments. Which scenario clearly develops the point/contribution you want to make? Shouldn't the other one occupy a smaller place (i.e. space) of the paper?)
		\end{itemize}
	\item Costly state verification
		\begin{itemize}
				\item Have you seen the work that Beshkar and Bond have done on what they call `contingent protection'?  I can't remember if they are drawing on the work of other people when they refer to the escape clause as `costly state verification.'  Their idea is that if you want to use the escape clause then this comes at the cost of providing information to the GATT panel to persuade them that you really have been hit by a shock.  This seems to provide an alternative way of separating genuine shocks from those created by lobby groups.  How does your approach relate to costly state verification?
		\end{itemize}
	\item Define $\ga$ before using it
	\item Take out repeats, especially of references to BS2005, rely to heavily on BS2005: In some parts the paper also seem repeating itself: you review Bagwell and Staiger (2005) in Section 2.2 but you provide another summary of the same paper in Section 3. Again, a revised structure may reduce the overlap, make the paper shorter, and be more to the point (and in general it seems to me that you seem to "dependent" on Bagwell and Staiger... (e.g. beginning of Sections 4 and 6).
	\item Scrub unnecessary references to weak bindings (since strong are now mostly gone)
	\item Perhaps make a bigger deal of the idea of $\ga$ decreasing function of $\tau$---move out of footnote
\end{enumerate}
